```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  width = 100,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

## Load packages

```{r}
library('car')
library('lsmeans')
library('plyr')
library('dplyr')
library('ggplot2')
library('knitr')
library('scales')
```

## Define some helper functions

```{r}
get_means <- function(x){
  list(means = as.data.frame(summary(x[[1]])), 
       diffs = as.data.frame(summary(x[[2]]))
  )
}

gg <- function(){
  list(
    geom_point(size = 4),
    geom_errorbar(limits, size = 1, width=0.2),
    theme_grey(base_size = 18)
  )
}

limits <- aes(ymax = lsmean + SE, ymin = lsmean - SE)

# Define function to get quick summary by factor variable
avg_by <- function(...){
  dat_df %>% 
    group_by(...) %>%
    summarise(rdm_mean = mean(rdm, na.rm = TRUE),
              rdm_sd = sd(rdm, na.rm = TRUE),
              rdm_n = length(na.omit(rdm))) %>%
    kable(format = "markdown")
}
```

## Summary of data

Prep data

```{r dataprep}
dat <- read.csv("dittes_data2.csv", stringsAsFactors = FALSE)
dat$date <- as.Date(paste(as.character(dat$year), "-01-01", sep = ""), "%Y-%m-%d")
dat <- dat[ !dat$habitat == 2, ] # drop habitat=32
dat$rdm <- dat$rdm*100 # multiply rdm by 100
dat$landform <- tolower(dat$landform)
```

Make a `dplyr` `tbl_df` object

```{r}
dat_df <- tbl_df(dat)
```

by transect

```{r results='asis'}
avg_by(transect)
```

by year

```{r results='asis'}
avg_by(year)
```

by soil type

```{r results='asis'}
avg_by(soil)
```

by slope_class

```{r results='asis'}
avg_by(slope_class)
```

by landform

```{r results='asis'}
avg_by(landform)
```

by habitat

```{r results='asis'}
avg_by(habitat)
```

All combinations of variables

```{r eval=FALSE}
avg_by(transect, year, soil, slope_class, landform, habitat) %>% data.frame
```


## Data analysis

All analyses are done with `rdm` (residual dry matter in lbs/sq acre) as the response variable. Plots use least square means for the y-variables since these take into account the statistical model that was used, wehreas raw data, although more meaningful, may be misleading. Confidence intervals on plots are 95% CI's.

### Prepare data a bit

```{r}
dat <- dat_df %>% filter(!is.na(rdm), !rdm < 2) %>% mutate(soil = as.factor(soil)) 
```

### Does RDM differ between soil types?

```{r anova_soils, results='asis'}
mod <- dat %>% lm(log10(rdm + 1) ~ soil, data=.) 
# mod %>% plot
mns <- mod %>% lsmeans(list(pairwise ~ soil), adjust = c("tukey")) %>% get_means
mod %>% Anova(type = "3") %>% kable(format = "markdown")
```

Plot by soil type

```{r}
mns$means %>% ggplot(aes(soil, lsmean)) + gg()
```

### Does RDM differ among years within each transect? 

```{r anova_transect, results='asis'}
mods <- dat %>% mutate(year = as.factor(year)) %>% group_by(transect) %>% do(model = lm(log10(rdm + 1) ~ year, data=.))
# mods$model[[1]] %>% plot
mns_s <- lapply(mods$model, function(x) x %>% lsmeans(list(pairwise ~ year), adjust = c("tukey")) %>% get_means)
for(i in seq_along(mods$model)){
  cat("\n")
  cat(sprintf("__%s__\n", mods$transect[i]))
  mods$model[[i]] %>% Anova(type = "3") %>% kable(format = "markdown")
}
```

Plot for each transect, among years

```{r}
rbind_all(Map(function(x,y) data.frame(transect=y, x), lapply(mns_s, "[[", "means"), unique(dat_df$transect))) %>%
  mutate(date = as.Date(paste0(year,"-01-01"), "%Y-%m-%d")) %>%
  ggplot(aes(date, lsmean)) + gg() + facet_wrap(~ transect) + scale_x_date(labels = date_format("%Y"))
```

### Power analysis (to figure out appropriate sample size)

> NOT DONE YET...
